---
title: "Forecasting with Two-Way ANOVA in R - When interaction is present"
author: "Anita Owens"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
---

## 1. Set up environment

```{r Load packages}
# Install pacman if needed
if (!require("pacman")) install.packages("pacman")

# load packages
pacman::p_load(pacman,
  tidyverse, openxlsx, ggpubr)
```

## 2. Import data

```{r Import games dataset}
#Dataset is in subfolder
(games <- read.xlsx("datasets/games.xlsx", sheet = 3))
```

```{r Make dataset long}
(games_long <- games %>% 
  pivot_longer(cols = starts_with("Price"), #columns to pivot to rows
               names_to = "price", #name the new column for the price variable
               values_to = "sales")) #the new sales column
```

```{r Re-level categorical variables}
games_long <- games_long %>% 
  mutate(advertising = factor(Advertising, levels = c("Low", "Medium", "High")),
  price = factor(price, levels = c("PriceLow", "PriceMedium", "PriceHigh"))) %>% 
  select(-Advertising)

str(games_long)
```


## 3. Visualize Data

```{r Visualize sales by advertising spend and price}
#Plot using ggpubr
ggline(games_long, x = "advertising", y = "sales", 
       add = c("mean_se", "jitter"),
       color = "price", palette = "jco",
       title = "Sales increase when ad spending increases",
       subtitle = "(but not when prices are high)",
       legend.title = "price levels"
       )
```

## 4. Two-way ANOVA model

The aov function models each combination of advertising and price on sales.

The f-values and p-values suggests that we reject the null hypothesis for advertising, price, and the interaction between advertising and price.

```{r Two-way ANOVA model}
two_way_aov <- aov(sales ~ advertising*price, data = games_long)

#The anova table
summary(two_way_aov)
```

How do we use this information from the aov table to forecast?

Since the p-value for the interaction (advertising\*price) is very small, we can forecast sales for any price and advertising combination equal to the mean of the observations.

```{r Equivalent to two-way aov but using lm() function, include = FALSE}
lm_model <- lm(sales ~  advertising*price, data = games_long)
summary(lm_model)
```

## 5. Model Diagnostics

```{r Model Diagnositic - Are ANOVA assumptions met}
#histogram with kernel density line
#hist(two_way_aov$residuals, prob = TRUE)
#lines(density(two_way_aov$residuals))

#However, when you have a small dataset it's better to visualize the residuals using qqplot.
qqnorm(two_way_aov$residuals)
qqline(two_way_aov$residuals)
#We have a few points quite far away from the line
```

## 6. Sales Forecast Table

```{r Forecast}
(forecast_tab <- games_long %>% 
  group_by(advertising, price) %>% 
  summarize(forecast_sales = round(mean(sales),2),
            std_dev = round(sd(sales),2)))

#Write to csv file
write.csv(forecast_tab,"datasets/games_forecast.csv", row.names = FALSE)
```

Because the p-value for the interaction is low, we have a significant interaction between advertising and price. Forecast if advertising is high and price is medium is 35.33

## 7. Final Summary

A key insight for marketing stakeholders, would be to cut back on advertising when the price is high.

Two-way ANOVA (with replication), if interaction effect **is significant**, then the predicted value is the value of the response variable (y) is equal to the mean of all observations having that combination of factor levels.


```{r Normality function - Check for normality in games dataset, include=FALSE}
normality_function <- function(y, x, stat_test){
  result <- tapply(y, x, stat_test)
  for(i in 1:length(result)){
    if(result[[i]][[2]] < 0.05){
      print("Not normal distribution")
    } else{
      print("Distributions are normal")
    }
  }
}

normality_function(games_long$sales, games_long$advertising, shapiro.test)
```

```{r Check for normality in games dataset - the long way, include=FALSE}
# Test normality across groups (Shapiro)
tapply(games_long$sales, games_long$advertising, FUN = shapiro.test)

tapply(games_long$sales, games_long$price, FUN = shapiro.test)
```



```{r Check homogeneity of variance in games dataset, include=FALSE}
# Check the homogeneity of variance (Bartlett)
bartlett.test(sales ~ advertising, data = games_long)
bartlett.test(sales ~ price, data = games_long)
#Both p-values are very large so homogeneity of variance is okay.
```